An Empirical Analysis of the Role of Energy in Economic Growth

### 2015-01-06 Articles/Article_EnerEcon_2014/EconGrowth3.Rnw
Added Csontos reference.

Some of the uncertainty in parameter estimates is explained by 
the mathematical structure of the CES model itself.%
  \footnote{
  Numbers in parentheses indicate rows in Table~\ref{tab:ces_sensitivity_table}.
  }
(1, 2)~Equations~\ref{eq:CES}--\ref{eq:CESekl} indicate that 
if $\delta_1 = 0$
or $\delta_1 = 1$,
$\rho_1$ cancels.
As a result, 
$sse$ becomes increasingly insensitive to $\rho_1$ (and, therefore, $\sigma_1$)
as $\delta_1 \rightarrow 0$ or $\delta_1 \rightarrow 1$.
Decreasing sensitivity leads to reduced precision in estimating $\sigma_1$.
%
(3)~If $\delta = 0$, $\delta_1$ and $\rho_1$ are eliminated.
Furthermore, with $\delta = 0$, $\rho$ cancels, 
and $sse$ becomes insensitive to $\rho$ and, therefore, $\sigma$.
Thus, as $\delta \rightarrow 0$, the precision with which 
$\delta_1$, $\sigma_1$, and $\sigma$ are estimated declines.
%
(4)~If $\delta = 1$, $\rho$ cancels, 
and $sse$ becomes insensitive to values of $\sigma$.
Given that both (a)~ the \texttt{cesEst} fitting algorithm
still uses a meaningless value for $\rho$ when $\delta \rightarrow 1$
and (b)~the ratio $\rho/\rho_1$ appears in the CES equation,
the value of $\rho_1$ will also become highly uncertain 
as $\delta \rightarrow 1$
due to the large uncertainty of $\rho$.
Essentially, $\rho_1$ compensates for imprecise values of $\rho$
on each successive resample,
taking on $\rho$'s uncertainty.
Thus, as $\delta \rightarrow 1$, 
the precision with which both $\sigma$ and $\sigma_1$ can be estimated declines.
%
(5)~If $\rho_1 = \infty$, $\sigma_1 = 0$, 
and the CES function (Equation~\ref{eq:CES}) simplifies to the Leontief function:
%
\begin{equation} \label{eq:ces_simplified_2_terms}
  y = \gamma A \min \left[ k, \, l \right] \; ,
\end{equation}
%
and $\delta_1$ is eliminated.
Thus, as $\sigma_1 \rightarrow 0$, estimation of $\delta_1$ becomes increasingly imprecise.
(6)~If $\rho = \infty$, $\sigma = 0$.
Using Equation~\ref{eq:CESkle} as an example, we find that 
%
\begin{equation} \label{eq:ces_simplifies_3_terms}
  y = \gamma A \min \left[ \delta_1 k^{-\rho_1} + (1-\delta_1) l^{-\rho_1}, \; e \right] \; ,
\end{equation}
%
and $\delta$ is eliminated.
Thus, as $\sigma \rightarrow 0$, estimation of $\delta$ becomes increasingly imprecise.

\begin{table} \caption{Implications of CES model structure 
                        on parameter precision.} \label{tab:ces_sensitivity_table} 
  \begin{center}
    \begin{tabular}{c c l c l} 
      \toprule
       & \multicolumn{2}{c}{Mathematical effect} & \multicolumn{2}{c}{Affect on precision} \\
      \midrule
% Row 1
      1 & $\delta_1 = 0$ & $\rho_1$ cancels & $\delta_1 \rightarrow 0$ & $\sigma_1$ imprecise \\ [2 mm]
% Row 2
      2 & $\delta_1 = 1$ & $\rho_1$ cancels & $\delta_1 \rightarrow 1$ & $\sigma_1$ imprecise \\ [2 mm]
% Row 3
      \multirow{3}{*}{3} 
        & \multirow{3}{*}{$\delta = 0$}   
                         & $\delta_1$ eliminated
                                            & \multirow{3}{*}{$\delta \rightarrow 0$}
                                                                       & $\delta_1$ imprecise \\ [-1 mm]
%
        &                & $\rho_1$ eliminated &                       & $\sigma_1$ imprecise \\ [-1 mm]
%
        &                & $\rho$\; cancels   
                                            &                          & $\sigma$\; imprecise \\ [2 mm]
% Row 4
      \multirow{2}{*}{4}
        & \multirow{2}{*}{$\delta = 1$}
                         & \multirow{2}{*}{$\rho$\; cancels}
                                            & \multirow{2}{*}{$\delta \rightarrow 1$}
                                                                       & $\sigma$\; imprecise   \\ [-1 mm] 
%
        &                &                  &                          & $\sigma_1$ imprecise    \\ [2 mm]
% Row 5
      \multirow{2}{*}{5}
        & \multirow{2}{*}{$\rho_1 = \infty$}
                         & $\sigma_1 = 0$   & \multirow{2}{*}{$\sigma_1 \rightarrow 0$}
                                                                       & \multirow{2}{*}{$\delta_1$ imprecise} 
                                                                                    \\ [-1 mm]
%
        &                & $\delta_1$ eliminated
                                            &                          &                      \\ [2 mm]
% Row 6
      \multirow{2}{*}{6}
        & \multirow{2}{*}{$\rho = \infty$}
                         & $\sigma = 0$     & \multirow{2}{*}{$\sigma \rightarrow 0$}
                                                                       & \multirow{2}{*}{$\delta$ imprecise} 
                                                                                    \\ [-1 mm]
%
        &                & $\delta$\; eliminated
                                            &                          &                      \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

>>>>

Some of the uncertainty in parameter estimates is explained by 
the mathematical structure of the CES equations.
Table~\ref{tab:ces_sensitivity_table} summarizes these observations.
Numbers in parentheses below correspond to row numbers in the table.
(1,~2)~Equations~\ref{eq:CES}--\ref{eq:CESekl} indicate that 
if $\delta_1 = 0$
or $\delta_1 = 1$,
$\rho_1$ cancels.
Consequently, 
$sse$ (Equation~\ref{eq:sse}) is insensitive to $\rho_1$ (and, therefore, $\sigma_1$).
As $\delta_1 \rightarrow 0$ or $\delta_1 \rightarrow 1$,
the precision with which $\sigma_1$ can be estimated declines.
%
(3)~If $\delta = 0$, $\delta_1$ and $\rho_1$ are eliminated.
Furthermore, with $\delta = 0$, $\rho$ cancels
and $sse$ becomes insensitive to $\rho$~($\sigma$).
Thus, as $\delta \rightarrow 0$, the precision with which 
$\delta_1$, $\sigma_1$, and $\sigma$ can be estimated declines.
%
(4)~If $\delta = 1$, $\rho$ cancels
and $sse$ becomes insensitive to values of $\sigma$.
Given that both 
(a)~ the \texttt{cesEst} fitting algorithm
still uses a meaningless value for $\rho$ when $\delta \rightarrow 1$ and 
(b)~the ratio $\rho/\rho_1$ appears in the CES equation,
the value of $\rho_1$ will also become imprecise
as $\delta \rightarrow 1$
due to the large uncertainty of $\rho$.
Essentially, $\rho_1$ compensates for imprecise values of $\rho$
on each successive resample,
thereby taking on $\rho$'s uncertainty.
Thus, as $\delta \rightarrow 1$, 
the precision with which both $\sigma$ and $\sigma_1$ can be estimated declines.
%
(5)~If $\rho_1 = \infty$, $\sigma_1 = 0$
and the CES function (Equation~\ref{eq:CES}) simplifies to the Leontief function
\citep{Csontos:1992aa}:
%
\begin{equation} \label{eq:ces_simplified_2_terms}
  y = \gamma A \min \left[ k, \, l \right] \; ,
\end{equation}
%
and $\delta_1$ is eliminated.
Thus, as $\sigma_1 \rightarrow 0$, 
the precision with which we can estimate $\delta_1$ declines.
%
(6)~If $\rho = \infty$, $\sigma = 0$.
Using Equation~\ref{eq:CESkle} as an example, we find that 
%
\begin{equation} \label{eq:ces_simplifies_3_terms}
  y = \gamma A \min \left[ \delta_1 k^{-\rho_1} + (1-\delta_1) l^{-\rho_1}, \; e \right] \; ,
\end{equation}
%
and $\delta$ is eliminated \citep{Csontos:1992aa}.
Thus, as $\sigma \rightarrow 0$, the precision with which $\delta$ can be estimated declines.

\begin{table} \caption{Implications of CES model structure 
                        on parameter precision.} \label{tab:ces_sensitivity_table} 
  \begin{center}
    \begin{tabular}{c c l c l} 
      \toprule
       & \multicolumn{2}{c}{Mathematical effect} & \multicolumn{2}{c}{Affect on precision} \\
      \midrule
% Row 1
      1 & $\delta_1 = 0$ & $\rho_1$ cancels & $\delta_1 \rightarrow 0$ & $\sigma_1$ imprecise \\ [2 mm]
% Row 2
      2 & $\delta_1 = 1$ & $\rho_1$ cancels & $\delta_1 \rightarrow 1$ & $\sigma_1$ imprecise \\ [2 mm]
% Row 3
      \multirow{3}{*}{3} 
        & \multirow{3}{*}{$\delta = 0$}   
                         & $\delta_1$ eliminated
                                            & \multirow{3}{*}{$\delta \rightarrow 0$}
                                                                       & $\delta_1$ imprecise \\ [-1 mm]
%
        &                & $\rho_1$ eliminated &                       & $\sigma_1$ imprecise \\ [-1 mm]
%
        &                & $\rho$\; cancels   
                                            &                          & $\sigma$\; imprecise \\ [2 mm]
% Row 4
      \multirow{2}{*}{4}
        & \multirow{2}{*}{$\delta = 1$}
                         & \multirow{2}{*}{$\rho$\; cancels}
                                            & \multirow{2}{*}{$\delta \rightarrow 1$}
                                                                       & $\sigma$\; imprecise   \\ [-1 mm] 
%
        &                &                  &                          & $\sigma_1$ imprecise    \\ [2 mm]
% Row 5
      \multirow{2}{*}{5}
        & \multirow{2}{*}{$\rho_1 = \infty$}
                         & $\sigma_1 = 0$   & \multirow{2}{*}{$\sigma_1 \rightarrow 0$}
                                                                       & \multirow{2}{*}{$\delta_1$ imprecise} 
                                                                                    \\ [-1 mm]
%
        &                & $\delta_1$ eliminated
                                            &                          &                      \\ [2 mm]
% Row 6
      \multirow{2}{*}{6}
        & \multirow{2}{*}{$\rho = \infty$}
                         & $\sigma = 0$     & \multirow{2}{*}{$\sigma \rightarrow 0$}
                                                                       & \multirow{2}{*}{$\delta$\; imprecise} 
                                                                                    \\ [-1 mm]
%
        &                & $\delta$\; eliminated
                                            &                          &                      \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}
